Log or not - Analysis of ECDC forecasts

### 2022-07-08 analysis/Analysis ECDC forecasts.Rmd
Update ECDC analasys and manuscript text

```{r}
scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
scores[, scale := factor(scale, levels = c("natural", "log"))]

@@

## Mean vs. sd of observed values

```{r}

# average number of cases and deaths
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value))

# negative binomial
# MASS::glm.nb(true_value ~ target_type, 
#        data = hub_data %>%
#          mutate(true_value = pmax(true_value, 0)), 
#        link = "identity")

@@

hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_minimal() + 
  scale_y_continuous(labels = label_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "Location")

ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 4)

# plot with mean vs. variance --------------------------------------------------
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(location, target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  # scale_x_continuous(labels = label_fn) +
  # scale_y_continuous(labels = label_fn) +
  labs(y = "Variance observed value", x = "Mean observed value") + 
  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
  theme_minimal()

@@

scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")

ggsave("output/figures/HUB-average-scores.png", width = 7, height = 3.5)
```

@@

## Scores across locations

--> could also do this plot with multiplicative scores, i.e. how much higher is a score than the lowest score observed

```{r}
# alternatively, we could think about just summarising across all models for these plots?

# make two separate plots to keep ordering. I.e. order locations according 
# to log scale to illustrate that the ordering changes for cases

get_order <- function(df, type = "Cases") {
  df |> 
    filter(scale == "natural", target_type == type) |>
    arrange(-interval_score) |>
    pull(location)
}

plot_df <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last")

order_cases <- get_order(plot_df)
order_deaths <- get_order(plot_df, type = "Deaths")

plot <- function(plot_df, order) {
  plot_df |>
    mutate(location = factor(location, levels = order)) |>
    ggplot(aes(y = interval_score, x = location)) + 
    geom_bar(stat = "identity") + 
    scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
    facet_wrap(~ target_type + scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "none") + 
    scale_y_continuous(labels = scale_fn) + 
    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
    labs(y = "Interval score", x = "Location")
}

cases <- plot_df |>
  filter(target_type == "Cases") |>
  plot(order = order_cases)

deaths <- plot_df |>
  filter(target_type == "Deaths") |>
  plot(order = order_deaths)

cases / 
  deaths

ggsave("output/figures/HUB-scores-locations.png", width = 7, height = 5)
```

*We see that the scores are much more evenly distributed across locations if we log the data*

@@

## Variance scores across locations

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(mean_wis = mean(interval_score), 
            sd_wis = sd(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = sd_wis, x = mean_wis)) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
  geom_point() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  # scale_x_continuous(labels = scale_fn) + 
  # scale_y_continuous(labels = scale_fn) + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(x = "Mean WIS", y = "Sd of WIS")

ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)

wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {

  plot_df |>
    ggplot(aes(y = wis, x = total)) + 
    geom_point() + 
    facet_wrap(~ target_type + scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Total number of observed values")
}

## Scores over time

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location %in% c("GB", "ES", "DE"),
         horizon == 2) |>
  rename(Location = location) |>
  group_by(target_type, scale, Location) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")

@@

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_minimal() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")

@@

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  labs(y = "Interval score", x = "Target type")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "GB",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "FR",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "ES",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_minimal() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
```

@@

>>>>

```{r}
scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
scores[, scale := factor(scale, levels = c("natural", "log"))]
data.table::setnames(scores, old = c("aem", "sharpness"), new = c("ae_median", "dispersion"))

@@

## =============================================================================
## FIGURE 4
## Mean vs. sd of observed values

```{r}
# average number of cases and deaths
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value))

@@

mean_observations <- hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value))

plot_means_obs <- mean_observations |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity", fill = "grey50") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_scoringutils() + 
  scale_y_continuous(labels = label_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "Location")

plot_df <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |> 
  full_join(mean_observations)

plot_mean_scores <- plot_df |>
  ggplot(aes(y = interval_score,
             x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(scale ~ target_type, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
  labs(y = "Interval score", x = "Location")

plot_means_obs / 
  plot_mean_scores + 
  plot_layout(heights = c(1, 2))

ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 8)

# This uses the ensemble, but alternatively, we could think about just summarising across all models for these plots? 
```

```{r}
# plot with mean vs. variance --------------------------------------------------
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(location, target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  # scale_x_continuous(labels = label_fn) +
  # scale_y_continuous(labels = label_fn) +
  labs(y = "Variance observed value", x = "Mean observed value") + 
  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
  theme_scoringutils() + 
  geom_smooth(alpha = 0.2, size = 0.2)

@@

scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")

ggsave("output/figures/HUB-average-scores.png", width = 7, height = 2.5)
```

@@

@@

## Scores over time
restrict to a few locations, e.g. GB, FR, ES (highest average scores there)

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location %in% c("GB", "ES", "DE"),
         horizon == 2) |>
  rename(Location = location) |>
  group_by(target_type, scale, Location) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")

@@

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_scoringutils() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")

@@

## Actual model rankings for single forecasts

```{r}
# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
# take one example
test_data <- hub_data |>
  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
         forecast_date == "2021-10-18", 
         target_type == "Cases", 
         location == "AT")

test_data |>
  score(metrics = c("interval_score")) |>
  summarise_scores()

test_data |>
  mutate(prediction = log(prediction), 
         true_value = log(true_value)) |>
  score(metrics = c("interval_score")) |>
  summarise_scores()
## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
# It seems this is because the trade-off between calibration and sharpness changes? 
# need to check in what way

rankings <- scores |>
  select(-c(coverage_deviation, bias, aem, sharpness, underprediction, overprediction)) |>
  group_by(location, forecast_date, horizon, target_type, scale) |>
  mutate(ranking = rank(interval_score) / n()) |>
   filter(forecast_date == "2021-10-18", 
         target_type == "Cases", 
         location == "AT")

factor_levels <-  rankings |>
  filter(horizon == 2, scale == "natural") |>
  group_by(model) |>
  summarise(ranking = mean(ranking)) |>
  arrange(-ranking) |>
  pull(model)

p_rankings <- rankings |>
  filter(horizon == 2) |>
  mutate(model = factor(model, levels = factor_levels)) |>
  ggplot(aes(y = model, x = ranking)) +
  ggdist::stat_halfeye() +
  facet_wrap(~ scale) +
  theme_scoringutils() +
  labs(y = "Model", x = "Normalised ranking")

rankings_diff <- rankings |>
  select(model, location, target_end_date, forecast_date, interval_score, scale, ranking) |>
  pivot_wider(names_from = scale, values_from = ranking, id_cols = -interval_score)

p_diff_ranking <- rankings_diff |>
  filter(horizon == 2) |>
  mutate(diff_ranking = natural - log) |>
  mutate(model = factor(model, levels = factor_levels)) |>
  ggplot(aes(y = model, x = diff_ranking)) +
  ggdist::stat_halfeye() +
  theme_scoringutils() +
  labs(y = "", x = "Differences in ranking") +
  theme(axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank())

p_rankings + p_diff_ranking + 
  plot_layout(widths = c(2, 1))

ggsave("output/figures/HUB-ranking-diffs.png", width = 9, height = 5)

rankings_diff |>
  group_by(horizon, target_type) |>
  summarise(cor = cor(natural, log))
```

## Look at pairwise comparisons

```{r}
summarised_pairwise <- scores |>
  filter(horizon == 2, 
         target_type == "Cases") |>
  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)

ranking_natural <- 
  summarised_pairwise |>
  filter(scale == "natural") |>
  arrange(-relative_skill) |>
  pull(model)

plot_rel_skill <- summarised_pairwise |>
  mutate(model = factor(model, levels = ranking_natural)) |>
  ggplot(aes(y = model, x = relative_skill)) +
  geom_bar(stat = "identity") +
  theme_scoringutils() +
  facet_wrap(~scale) +
  labs(y = "Model", x = "Relative skill")

diffs <- summarised_pairwise |>
  mutate(model = factor(model, levels = ranking_natural)) |>
  select(model, relative_skill, scale) |>
  pivot_wider(names_from = scale, values_from = relative_skill) |>
  mutate(model = factor(model, levels = ranking_natural), 
         difference = natural - log)

plot_diffs <- diffs |>
  mutate(title = "difference") |>
  ggplot(aes(y = model, x = difference)) +
  geom_bar(stat = "identity") +
  theme_scoringutils() +
  labs(y = "Model", x = "Difference in relative skill") +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  facet_wrap(~title)

plot_distr_scores <- scores |>
  filter(horizon == 2) |>
  mutate(model = factor(model, levels = ranking_natural)) |>
  ggplot(aes(y = model, x = interval_score)) +
  ggdist::stat_pointinterval() +
  facet_wrap(~ scale, scale = "free_x") +
  theme_scoringutils() +
  scale_x_log10(labels = label_fn) +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  labs(y = "Model", x = "WIS")

plot_rel_wis <- scores |>
  filter(horizon == 2, 
         target_type == "Cases") |>
  mutate(model = factor(model, levels = ranking_natural)) |>
  summarise_scores(by = c("model", "scale")) |>
  plot_wis(relative_contributions = FALSE) +
    theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  scale_x_continuous(labels = label_fn) +
  facet_wrap(~scale, scales = "free_x")

plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores +
  plot_layout(widths = c(2, 1, 2)) &
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

ggsave("output/figures/HUB-pairwise-comparisons.png", width = 9, height = 5)

#Issues: this currently ignores that some models have only made forecasts for one location. They should usually be excluded, but maybe not necessary here as we're really more comparing the results for natural and log rather than the models against each other.

cor(diffs$log, diffs$natural)

# "CovidMetrics-epiBATS" only has forecasts in DE
# "ICM-agentModel" only in Poland
# "UVA is everywhere
hub_data |>
  filter(model == "UVA-Ensemble") |> check_forecasts()

plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases") {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

  locations <- unique(filtered_hub$location)

  filtered_scores <- scores |>
    filter(model == forecast_model, 
           horizon %in% h, 
           location %in% locations,
           target_type == type)

  plot_natural  <- filtered_hub |>  
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    theme(legend.position = "bottom")

  plot_log <- filtered_hub |>
    mutate(prediction = log(prediction + 1), 
           true_value = log(true_value + 1)) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    theme(legend.position = "bottom")

  scores_natural <- filtered_scores |>
    filter(scale == "natural") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    scale_x_continuous(labels = label_fn)

  scores_log <- filtered_scores |>
    filter(scale == "log") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE)

  (plot_natural + plot_log) /
    (scores_natural + scores_log) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}

ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")

epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")

BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")

ggsave(filename = "output/figures/HUB-model-ICM.png", width = 7, height = 4.5, plot = ICM)
ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 7, height = 4.5, plot = epiBATS)
ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 7, height = 4.5, plot = BSLCoV)

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  labs(y = "Interval score", x = "Target type")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "GB",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "FR",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "ES",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

@@

```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_scoringutils() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
```

@@

## Code parking lot

## Variance of scores across locations

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(mean_wis = mean(interval_score), 
            sd_wis = sd(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = sd_wis, x = mean_wis)) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
  geom_point() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  # scale_x_continuous(labels = scale_fn) + 
  # scale_y_continuous(labels = scale_fn) + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(x = "Mean WIS", y = "Sd of WIS")

ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
```

```{r}
# negative binomial
# MASS::glm.nb(true_value ~ target_type, 
#        data = hub_data %>%
#          mutate(true_value = pmax(true_value, 0)), 
#        link = "identity")
```

## WIS vs. the totsl number of observed values
probably not needed anymore, as we have similar things

```{r}
wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {

  plot_df |>
    ggplot(aes(y = wis, x = total)) + 
    geom_point() + 
    facet_wrap(~ target_type + scale, scale = "free") + 
    theme_scoringutils() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Total number of observed values")
}
